---
title: "improved leverage"
author: "zhu"
date: "23/12/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
```
```{r}
source('utils.R')
```
```{r data}
set.seed(999)
####pipeline#####
i=2#treatment
j=6#time
r=4#replicate
time=c(0,1,2,3,4,5)#time points
#y1 <- 0.01*time+0.01*time^2+0.01*time^3+0.01*time^4
ctrl <- 0.4*time
tre1 <- -0.4*time
tre2 <- -0.2*((time-2.5)^2)+0.5
tre3 <- c(0,0.3,1,0.4,0.6,2)
notrend <- rep(0,6)
y1 <- c(notrend)
y12 <- c(tre1)

y2 <- c(notrend)
y22 <- c(tre1)

y3 <- c(notrend)
y32 <- c(tre1)
#y12 <- 0.01*time^2

#y2 <- -time

#y3 <- c(-2,-1,0,0,1,2)
# y1 <- c(0,1,2,3,4,5)
#y32 <- 0.01*time^2
#y3 <- -0.01*time^3

#total of 100 metabolites(columns)
#total of 2*5*4=40 rows
p1 <- 15
p2 <- 15
p3 <- 1
neg <- 2000
simulation <- generate_data.nocorrelation(p1,p2,p3,neg)
df.final <- simulation$df
groups <- simulation$groups
```

##Data

### there are 2 treatments, 6 time points and 4 replicates. replicates were drawn as independent observations from normal distribution N(0,0.2). The simulation dataset consists of 48 rows(samples) and 81 columns(31 differential expressed variable and 50 flat variables).

###example of differentially expressed variable is shown below.
```{r plot raw data metabolite one}
trend.toplot <- data.frame(replicate=rep(1:(i*j),each=r),time=rep(time,each=r),treatment=factor(rep(1:i,each=j*r)),metabolite=df.final[1,])
ggplot(trend.toplot,aes(x=time,y=metabolite,color=treatment))+
  geom_point()+
  stat_summary(fun.y = mean,geom = 'line')+
  labs(title = 'differentially expressed variable')
```

##Original leverage

```{r original asca-gene design matrix and fit, message=F, warning=F}
#design matrix
mx <- asca.design.matrix(i,j,r,time)
#fit model
type=1
Fac=c(1,1,1,1)
asca.fit <- ASCA.2f(X = t(df.final),Designa = mx$j, Designb = mx$i,type = type, Fac = Fac,showvar = F,showscree = T)
```

```{r leverage and spe plot}
plot.leverage_spe_original(df.final,asca.fit,groups,R=1,No.sample = 50)
```

###As expected, with 1 PC selected for submodel treatment and submodel interaction, leverage cutoff line set excluded all of the variabes to be significant.

##Improved leverage

###As discussed previously, this is due to the loadings were normalized to unit vector during calculation. Alternatively, we could normalize score to unit vector. Modification was done in the asca-gene algotirhm to achieve this. The formulus as well as the code is shown below:

```{r PCA ld unormed,eval=F, echo=T}
PCA.GENES.unorm_loading<-function(X)
{
  #PCA.GENES is very useful to obtain principal components to a matrix that has more variables than individuals. 
  #R can not apply princomp is such case and when there are a lot of variables eigen(t(X)%*%X) can not be computed.
  
  #X is a matrix that has on columns the genes considered as variables in the PCA analysis.
  #First we center the matrix by columns (Xoff) and then we obtain the eigenvalues and the eigenvectors of the matrix Xoff%*%t(Xoff) and we #use the equivalences between the loadings and scores to obtain the solution
  
  
  n<-ncol(X)
  p<-nrow(X)
  offset<-apply(X,2,mean)
  Xoff<-X-(cbind(matrix(1,p,1))%*%rbind(offset))
  
  eigen<-eigen(Xoff%*%t(Xoff)/(p-1))#calculates eigen value and unit eigen vector of c*cT
  var<-cbind(eigen$values/sum(eigen$values),cumsum(eigen$values/sum(eigen$values)))
  
  scores.normed<-eigen$vectors
  loading.unormed<-t(Xoff)%*%scores.normed
  

  
  output<-list(eigen,var,scores.normed,loading.unormed)
  names(output)<-c("eigen","var.exp","scores","loadings.unormed")
  output
}
```

```{r unormed lds,results=F,fig.height=5,fig.width=10}
asca.fit.normedld <- ASCA.2f_leverage(X = t(df.final),Designa = mx$j, Designb = mx$i,type = type, Fac = Fac,showvar = F,showscree = F)
plot.submodels_loading(asca.fit.normedld,groups,title="unormed loading for submodel b.ab")
```


```{r improved leverage,results=F}
asca.fitted <- fitted(df.final,asca.fit,R=1,groups=groups,which_leverage = ASCA.2f_leverage)
leverage_spe <- plot.leverage_spe(asca.fitted,0.2)
#leverage_spe$`metabolites>leverage_lim` %notin% groups[1:(p1+p2+p3)]
confusion_matrix <- stats(asca.fitted$stats_for_plot)
```

###the new loading and leverage and spe plot are shown above. simulation data are correctly indentified as designed.